---
title: "Project test observations onto existing topolgy"
author: 'Trang Le'
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project test observations onto existing topolgy}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

The {tdapseudotime} :package: contains functions to implement the temporal phenotyping via topological data analysis.

You can install the development version of tdapseudotime from GitHub with remotes:
```{r, eval=FALSE}
# install.packages('remotes') # uncomment to install devtools
remotes::install_github('trang1618/tdapseudotime')
```

```{r setup, message=FALSE, warning=FALSE}
library(tdapseudotime)
library(dplyr)
library(tidyr)
library(ggplot2)

set.seed(1234)
```

```{r echo=FALSE}
theme_set(
  theme_bw() + 
    theme(
      panel.grid.minor = element_blank(),
      legend.title = element_blank()
    ))
```

## Preprocess

Parameters:
```{r}
my_test <- 'test-data.csv'
data("centroids", package = 'tdapseudotime')
```

Color palette for enrichment: blue > green > yellow > orange > red
```{r}
# RColorBrewer::brewer.pal(length(unique(CommunityCluster$membership)), "Set1")
my_colors <- c("#00A3DD", "#60C659", "#FFBC21", "#FF7F1E", "#EF2B2D")
```

## Read in data file
and load the data exported from training, `centroids.rda`, which contains 
`f_graph` (igraph object), `out_trajectories`, and the `centroids` data frame.

```{r}
test_data <- read.csv(my_test, header = TRUE, colClasses = "character") %>% 
  rename('covid_id' = patient_num,
         'time' = days_since_admission)
non_lab_value_names <- c('id', 'covid_id', 'day')
lab_value_names <- setdiff(names(test_data), non_lab_value_names)


```

## Process data

```{r}
processed_test <- test_data %>% 
  group_by(covid_id) %>% 
  ungroup() %>% 
  mutate(id = as.integer(id),
         time = as.numeric(time)) %>% 
  arrange(covid_id, time) %>% 
  mutate_at(dplyr::all_of(lab_value_names), as.numeric) %>% 
  arrange(as.numeric(covid_id), time) %>% 
  {.}
```

```{r}
projection_mat <- cosine_sim_func(
  processed_test %>% select(- c('id', 'covid_id')) %>% as.matrix(), 
  as.matrix(select(centroids, - node)))

id_to_node <- data.frame(
  id = processed_test$id,
  node = apply(projection_mat, 1, which.max)
)
```

```{r}
mapped_test <- processed_test %>% 
  select(- any_of(lab_value_names)) %>% 
  left_join(id_to_node, by = 'id') %>% 
  arrange(id)

out_test_list <- compute_similarity(mapped_test, f_graph, out_trajectories)
similarity_test <- out_test_list[[1]]
id_node_cluster_test <- out_test_list[[2]]
most_similar_traj_test <- similarity_test %>%
  group_by(covid_id) %>%
  slice(which.max(SJ)) # use Jaccard similarity

head(most_similar_traj_test, 10)
test_out <- most_similar_traj_test %>% select(covid_id, clusterTraj)

table(test_out$clusterTraj)
```

## Visualizations

```{r}
plot_dat <- processed_test %>% 
  left_join(id_node_cluster_test %>% distinct(covid_id, id, cluster), 
            by = c('id', 'covid_id'))

plot_dat %>% 
  ggplot(aes(x = cluster, y = time, fill = cluster)) +
  geom_boxplot(alpha = 0.8) + 
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) +
  theme(legend.position = "none",
        plot.title = element_text(size = 8, hjust = 0.5))

plot_dat %>% 
  select(cluster, all_of(lab_value_names)) %>% 
  pivot_longer(- cluster, names_to = 'Lab', values_to = 'lab_value') %>% 
  ggplot(aes(x = cluster, y = lab_value, fill = cluster)) +
  geom_boxplot(alpha = 0.8) + 
  labs(x = NULL, y = NULL) +
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) +
  theme(legend.position = "none") +
  facet_wrap(~ Lab, scales = 'free_y')

processed_data_traj <- processed_test %>% 
  left_join(most_similar_traj_test, by = c("covid_id")) %>% 
  mutate(clusterTraj = as.factor(clusterTraj), time) %>%
  select(time, clusterTraj, all_of(lab_value_names)) %>% 
  distinct()
```

```{r fig.width=12, fig.height=8}
processed_data_traj  %>% 
  pivot_longer(- c(time, clusterTraj), 
               names_to = 'Lab', values_to = 'lab_value') %>% 
  ggplot(aes(time, lab_value, colour = clusterTraj, 
             group = clusterTraj, fill = clusterTraj)) +
  geom_smooth(method = "loess") +
  theme(legend.position = "none") +
  facet_wrap(~ Lab, scales = 'free_y')
```
