---
title: "Train: Establish topology"
author: 'Trang Le'
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Train: Establish topology}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

If your site is one of the "train" sites, please use this workflow to establish the topology and trajectories using the observations at your site.

## Setting up

```{r setup, message=FALSE, warning=FALSE}
library(tdapseudotime)
library(mice)
library(dplyr)
library(tidyr)
library(ggplot2)

set.seed(1234)
```

```{r echo=FALSE}
theme_set(
  theme_bw() + 
    theme(
      panel.grid.minor = element_blank(),
      legend.title = element_blank()
    ))
```

## Preprocess

Parameters:
```{r}
my_file <- 'train-data.csv'

# TDAMapper parameters (see grid search below)
n_intervals <- 6
p_overlaps <- 60
n_clusts <- 8
```

Color palette for enrichment: blue > green > yellow > orange > red
```{r}
# RColorBrewer::brewer.pal(length(unique(CommunityCluster$membership)), "Set1")
my_colors <- c("#00A3DD", "#60C659", "#FFBC21", "#FF7F1E", "#EF2B2D")
```

## Read in data file
Create a timeline for each subject (being time zero the first observation).
Extract lab values to be used for TDA distance matrix.

Check format:
- first col: row id "id" (can be any value, i.e. unique row number)
- second col: pts id "covid_id"
- third col: date "day"

*Need to check whether rowid is important*

```{r}
FupData <- read.csv(my_file, header = TRUE) %>% 
  widen_i2b2()

non_lab_value_names <- c('id', 'covid_id', 'CardiacTroponinHighSensitivity')
lab_value_names <- setdiff(names(FupData), non_lab_value_names)
processed_data <- mutate_at(FupData, dplyr::all_of(lab_value_names), as.numeric)
dim(processed_data)
```

```{r warning=FALSE, results='hide', message=FALSE}
mm <- 5
lab_values_mat <- processed_data[, lab_value_names] %>% 
  mice(m = mm, maxit = 25, meth = 'pmm', seed = 500) %>% # imputation
  # complete(1) %>%
  # scale() %>% # prepare for cosine similarity calculation
  {.}

lab_values_mat <- lapply(1:mm, function(x) complete(lab_values_mat, x)) %>%
  do.call(rbind, .) %>%
  as.matrix() %>%
  scale()
```

## Perform grid search to select a combination of hyperparameters
The code in this section is optional.

```{r fig.width=9}
# intervals_seq <- seq(3, 11, 1)
# overlaps <- 60
# clusts <- 8
# graph_grid <- list()
# 
# for (ii in seq_along(intervals_seq)){
#   print(ii)
#   intervals <- intervals_seq[ii]
#   f_sim_map <- map_tda(lab_values_mat,
#                        num_intervals = c(intervals, intervals),
#                        percent_overlap = overlaps,
#                        num_bins_when_clustering = clusts)
#   f_graph <- igraph::graph.adjacency(f_sim_map$adjacency, mode = "undirected")
#   
#   graph_grid[[ii]] <- c(intervals = intervals, overlaps = overlaps,
#                         clust_bins = clusts,
#                         get_graph_properties(f_graph, f_sim_map))
# }
# graph_grid <- do.call("rbind", graph_grid) %>% 
#   as.data.frame()
# 
# graph_grid %>% 
#   select(- overlaps, - clust_bins) %>% 
#   pivot_longer(-intervals) %>% 
#   ggplot(aes(x = intervals, y = value)) +
#   geom_point() +
#   scale_x_continuous(breaks = intervals_seq) +
#   facet_wrap(~ name, scales = 'free_y')

```

```{r eval=FALSE, include=FALSE}
# intervals_seq <- seq(3, 11, 1)
# # overlaps_seq <- seq(50, 100, 10)
# # clust_bins <- c(6, 8, 10, 12, 14)
# overlaps <- 50
# clusts <- 6
# graph_grid <- list()
# i <- 0
# for (ii in seq_along(intervals_seq)){
#   # for (p in seq_along(overlaps_seq)){
#     # for (b in seq_along(clust_bins)){
#       i <- i + 1
#       print(ii)
#       intervals <- intervals_seq[ii]
#       # overlaps <- overlaps_seq[p]
#       # clusts <- clust_bins[b]
# 
#       f_sim_map <- map_tda(lab_values_mat,
#                      num_intervals = c(intervals, intervals),
#                      percent_overlap = overlaps,
#                      num_bins_when_clustering = clusts)
#       f_graph <- igraph::graph.adjacency(f_sim_map$adjacency, mode = "undirected")
# 
#       graph_grid[[i]] <- c(intervals = intervals, overlaps = overlaps,
#                            clust_bins = clusts,
#                            get_graph_properties(f_graph, f_sim_map))
#     # }
#   # }
# }
# graph_grid <- do.call("rbind", graph_grid) 
# graph_grid %>% 
#   as.data.frame() %>% 
#   filter(n_comps == 1, clust_bins <= 10) %>% 
#   distinct(intervals, n_comps, n_nodes, median_degree, edge_density, clique_length, diameter, mean_distance, .keep_all = T)
```

We chose the hyperparameters to be 6 to be the number of intervals for TDAMapper.
```{r}
# graph_grid %>% 
#   # as.data.frame() %>% 
#   filter(intervals == 6)
```


## Run TDA Mapper
```{r}
f_sim_map <- map_tda(lab_values_mat,
                     num_intervals = c(n_intervals, n_intervals),
                     percent_overlap = p_overlaps,
                     num_bins_when_clustering = n_clusts)
f_graph <- make_tda_graph(
  f_sim_map, 
  data = processed_data, 
  enrich_var = 'time', # enrich topology by time for now
  color_method = 'clust_color',
  my_colors = my_colors
)
```

## Create a minimum spanning tree
and compute trajectories and assign observations to nodes in network
### THIS STEP NEED MANUAL REVIEW 
Check the MST plot and TIME boxplots

```{r}
out_trajectories <- find_trajectories(processed_data, f_sim_map, f_graph)
out_list <- compute_similarity(processed_data, f_graph$node_color, out_trajectories, f_sim_map)
```

```{r out.width=12}
similarity_df <- out_list[[1]]
id_node_cluster <- out_list[[2]]
most_similar_traj <- similarity_df %>%
  group_by(covid_id) %>%
  slice(which.max(SJ)) # use Jaccard similarity

head(most_similar_traj, 10)

centroids <- processed_data[, c('id', lab_value_names)] %>% 
  left_join(id_node_cluster[, c('id', 'node')], by = 'id') %>% 
  group_by(node) %>% 
  summarise(across(.fns = median), .groups = 'drop') %>% 
  select(-id)

node_color <- f_graph$node_color
# save(centroids, node_color, out_trajectories, file = '../data/centroids.rda')
```

## Write output
```{r}
data_out <- most_similar_traj %>% select(covid_id, clusterTraj)
table(data_out$clusterTraj)
```

## Visualizations

```{r fig.width=10}
plot_dat <- processed_data %>% 
  left_join(id_node_cluster %>% distinct(covid_id, id, cluster), 
            by = c('id', 'covid_id'))

plot_dat %>% 
  ggplot(aes(x = cluster, y = time, fill = cluster)) +
  geom_boxplot(alpha = 0.8) + 
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) +
  theme(legend.position = "none",
        plot.title = element_text(size = 8, hjust = 0.5))

plot_dat %>% 
  select(cluster, all_of(lab_value_names)) %>% 
  pivot_longer(- cluster, names_to = 'Lab', values_to = 'lab_value') %>% 
  ggplot(aes(x = cluster, y = lab_value, fill = cluster)) +
  geom_boxplot(alpha = 0.8) + 
  labs(x = NULL, y = NULL) +
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) +
  theme(legend.position = "none") +
  facet_wrap(~ Lab, scales = 'free_y')

processed_data_traj <- processed_data %>% 
  left_join(most_similar_traj, by = c("covid_id")) %>% 
  mutate(clusterTraj = as.factor(clusterTraj), time) %>%
  select(time, clusterTraj, all_of(lab_value_names)) %>% 
  distinct()
```

```{r fig.width=12, fig.height=8, warning=FALSE, message=FALSE}
processed_data_traj  %>% 
  pivot_longer(- c(time, clusterTraj), 
               names_to = 'Lab', values_to = 'lab_value') %>% 
  ggplot(aes(time, lab_value, colour = clusterTraj, 
             group = clusterTraj, fill = clusterTraj)) +
  geom_smooth(method = "loess") +
  theme(legend.position = "none") +
  facet_wrap(~ Lab, scales = 'free_y')
```

