---
title: "Establish topology"
author: 'Trang Le'
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Establish topology}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

If your site is one of the "train" sites, please use this workflow to establish the topology and trajectories using the observations at your site.

## Setting up

```{r setup, message=FALSE, warning=FALSE}
library(tdapseudotime)
library(mice)
library(dplyr)
library(tidyr)
library(ggplot2)

set.seed(1234)
```

```{r echo=FALSE}
theme_set(
  theme_bw() + 
    theme(
      panel.grid.minor = element_blank(),
      legend.title = element_blank()
    ))
```

## Preprocess

Parameters:
```{r}
my_file <- 'train-data.csv'

# TDA hyperparameter grid
intervals_seq <- seq(5, 10, 1)
overlaps_seq <- seq(50, 60, 10)
clust_bins <- c(6, 8, 10)
ii <- 2
p <- 2
b <- 2
num_intervals <- c(intervals_seq[ii], intervals_seq[ii])
percent_overlap <- overlaps_seq[p]
num_bins_when_clustering <- clust_bins[b]
```

Color palette for enrichment: blue > green > yellow > orange > red
```{r}
# RColorBrewer::brewer.pal(length(unique(CommunityCluster$membership)), "Set1")
my_colors <- c("#00A3DD", "#60C659", "#FFBC21", "#FF7F1E", "#EF2B2D")
```

## Read in data file
Create a timeline for each subject (being time zero the first observation).
Extract lab values to be used for TDA distance matrix.

Check format:
- first col: row id "id" (can be any value, i.e. unique row number)
- second col: pts id "covid_id"
- third col: date "day"

*Need to check whether rowid is important*

```{r}
FupData <- read.csv(my_file, header = TRUE, colClasses = "character")
non_lab_value_names <- c('id', 'covid_id', 'day')
lab_value_names <- setdiff(names(FupData), non_lab_value_names)

processed_data <- FupData %>% 
  mutate(day = as.Date(day, format = '%Y-%m-%d')) %>% 
  group_by(covid_id) %>% 
  mutate(first_date = min(day)) %>% 
  ungroup() %>% 
  mutate(id = as.integer(id),
         time = as.numeric(day - first_date, units = 'days')) %>% 
  arrange(covid_id, time) %>% 
  mutate_at(dplyr::all_of(lab_value_names), as.numeric)
```

```{r warning=FALSE, results='hide', message=FALSE}
lab_values_mat <- processed_data[, lab_value_names] %>% 
  mice(m = 5, maxit = 50, meth = 'pmm', seed = 500) %>% # imputation
  complete(1) %>% 
  scale() # prepare for cosine similarity calculation
```

## Run TDA Mapper
```{r}
f_sim_map <- map_tda(lab_values_mat,
                     num_intervals = c(intervals_seq[ii], intervals_seq[ii]),
                     percent_overlap = overlaps_seq[p],
                     num_bins_when_clustering = clust_bins[b])
f_graph <- make_tda_graph(
  f_sim_map, 
  data = processed_data, 
  enrich_var = 'time', # enrich topology by time for now
  color_method = 'clust_color',
  my_colors = my_colors
)
```

```{r include=FALSE, eval=FALSE}
# intervals_seq <- seq(5, 10, 1)
# overlaps_seq <- seq(50, 60, 10)
# clust_bins <- c(6, 8, 10)
# graph_grid <- list()
# i <- 0
# for (ii in seq_along(intervals_seq)){
#   for (p in seq_along(overlaps_seq)){
#     for (b in seq_along(clust_bins)){
#       i <- i + 1
#       intervals <- intervals_seq[ii]
#       overlaps <- overlaps_seq[p]
#       clusts <- clust_bins[b]
#       
#       f_sim_map <- map_tda(lab_values_mat,
#                      num_intervals = c(intervals, intervals),
#                      percent_overlap = overlaps,
#                      num_bins_when_clustering = clusts)
#       f_graph <- graph.adjacency(f_sim_map$adjacency, mode = "undirected")
#       
#       graph_grid[[i]] <- c(intervals = intervals, overlaps = overlaps, 
#                            clust_bins = clusts, 
#                            get_graph_properties(f_graph))
#     }
#   }
# }
# graph_grid <- do.call("rbind", graph_grid)

```

## Create a minimum spanning tree
and compute trajectories and assign observations to nodes in network
### THIS STEP NEED MANUAL REVIEW 
Check the MST plot and TIME boxplots

```{r}
out_trajectories <- find_trajectories(processed_data, f_sim_map, f_graph)
out_list <- compute_similarity(processed_data, f_graph, out_trajectories, f_sim_map)
```

```{r out.width=12}
similarity_df <- out_list[[1]]
id_node_cluster <- out_list[[2]]
most_similar_traj <- similarity_df %>%
  group_by(covid_id) %>%
  slice(which.max(SJ)) # use Jaccard similarity

head(most_similar_traj, 10)

centroids <- processed_data[, c('id', 'time', lab_value_names)] %>% 
  left_join(id_node_cluster[, c('id', 'node')], by = 'id') %>% 
  group_by(node) %>% 
  summarise(across(.fns = mean), .groups = 'drop') %>% 
  select(-id)

# save(centroids, f_graph, out_trajectories, file = '../data/centroids.rda')
```

## Write output
```{r}
data_out <- most_similar_traj %>% select(covid_id, clusterTraj)
table(data_out$clusterTraj)

```

## Visualizations

```{r}
plot_dat <- processed_data %>% 
  left_join(id_node_cluster %>% distinct(covid_id, id, cluster), 
            by = c('id', 'covid_id'))

plot_dat %>% 
  ggplot(aes(x = cluster, y = time, fill = cluster)) +
  geom_boxplot(alpha = 0.8) + 
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) +
  theme(legend.position = "none",
        plot.title = element_text(size = 8, hjust = 0.5))

plot_dat %>% 
  select(cluster, all_of(lab_value_names)) %>% 
  pivot_longer(- cluster, names_to = 'Lab', values_to = 'lab_value') %>% 
  ggplot(aes(x = cluster, y = lab_value, fill = cluster)) +
  geom_boxplot(alpha = 0.8) + 
  labs(x = NULL, y = NULL) +
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) +
  theme(legend.position = "none") +
  facet_wrap(~ Lab, scales = 'free_y')

processed_data_traj <- processed_data %>% 
  left_join(most_similar_traj, by = c("covid_id")) %>% 
  mutate(clusterTraj = as.factor(clusterTraj), time) %>%
  select(time, clusterTraj, all_of(lab_value_names)) %>% 
  distinct()
```

```{r fig.width=12, fig.height=8}
processed_data_traj  %>% 
  pivot_longer(- c(time, clusterTraj), 
               names_to = 'Lab', values_to = 'lab_value') %>% 
  ggplot(aes(time, lab_value, colour = clusterTraj, 
             group = clusterTraj, fill = clusterTraj)) +
  scale_fill_manual(values = f_graph$pal$color) + 
  scale_color_manual(values = f_graph$pal$color) + 
  geom_smooth(method = "loess") +
  theme(legend.position = "none") +
  facet_wrap(~ Lab, scales = 'free_y')
```



